\documentclass[11pt]{amsart}

\usepackage{mathptmx}
\usepackage{a4wide}
\usepackage{paralist}
\usepackage{url}
\usepackage{bbm}
\usepackage{fancyvrb}
\usepackage{xcolor}

\newcommand{\cA}{\mathcal{A}}
\newcommand{\cS}{\mathcal{S}}
\DeclareMathOperator{\conv}{conv}
\DeclareMathOperator{\New}{New}
\DeclareMathOperator{\area}{area}
\newcommand{\RR}{\mathbbm{R}}
\newcommand{\CC}{\mathbbm{C}}

\newcommand{\defn}[1]{{\color{blue}#1}}
\newcommand{\alert}[1]{\textbf{\color{red}#1}}

\DefineShortVerb{\;}
\renewcommand{\FancyVerbFormatCom}{\color{green!30!black}}

\begin{document}
\begin{center}
\textbf{\sffamily
   Discrete and Algorithmic Geometry }

\medskip
   Julian Pfeifle,
   UPC, 2019
\end{center}

\bigskip

\begin{center}
  \textbf{\sffamily Sheet 1}

  \bigskip\bigskip
Due on Tuesday, November 19, 2019
\end{center}

\bigskip


\begin{quote}\small
  To submit your solutions to these exercises,
  \begin{itemize}[$\quad\triangleright$]
  \item create a new branch ;your-awesome-team-name-sheet-1;,
  \item create a subdirectory ;exercises/sheet1/your-awesome-team-name/;,
  \item and put your solutions to the exercises into a ;.pdf; file into that directory.
  \item Now encrypt this ;.pdf; using ;julian.pfeifle@upc.edu.public.gpg.key;, and
  \item ;add;, ;commit; and ;push; \alert{only this encrypted pdf, not the original} ;.tex;
  \item and create a pull request.
  \end{itemize}
  You will be graded collectively on these exercises, and individually in the final exam.

  \alert{Exercises not submitted via this mechanism will not be graded.}
\end{quote}

\bigskip

Let $([n],\mathcal I)$ be a matroid on the ground set $[n]=\{1,2,\dots,n\}$ with independent sets~$\{I:I\in\mathcal I\}$.

\begin{itemize}[$\triangleright$]
\item
  For any proper subset $S\subset[n]$, the \defn{deletion}
  $M\setminus S$ is the matroid on the ground set $[n]\setminus S$ whose
  independent sets are $\{I\subset[n]\setminus S : I\in\mathcal I\}$.

\item
  The \defn{dual matroid} $M^\star$ of~$M$ is the matroid on $[n]$ where $I$~is a basis iff $[n]\setminus I$ is a basis~of~$M$.

\item
  If $S\subset[n]$, then the \defn{contraction} of $M$ with respect to $S$ is $M/S = (M^\star\setminus S)^\star$.

\item
  Let $G$ be a graph whose edges are labeled by $[n]$.
  The bases of the \defn{graphical matroid} $M_G$ are the sets of edges corresponding to spanning trees of~$G$.
\end{itemize}

\bigskip
\begin{enumerate}
\item True or false?
  \begin{enumerate}
  \item This notion of contraction agrees with the notion of
    contraction in graph theory.
  \item $M_{G^\star} = (M_G)^\star$, if $G$ is a planar graph and $G^\star$ its dual planar graph.
  \end{enumerate}
  
  \bigskip
\item Prove that if a matroid $M$ is realizable over a ground field $\mathbbm k$, then the dual matroid~$M^\star$ is also realizable over~$\mathbbm k$.
  {\footnotesize\color{green!30!black} [\emph{Hint.} Suppose that $M$ has rank~$d$ and $n$~elements.
  After a change of basis, $M$~can be realized by the $d\times n$ matrix $A=[I|B]$, where $I$~is the $d\times d$ identity matrix, and $B$~has size $d\times(n-d)$.
  Now find a matrix that realizes~$M^\star$.]
}

\bigskip
\item
  Consider the matroid $M$ realized by the columns of the matrix
  \[
    \begin{bmatrix}
      0 & 1 & 1 & 0 & 1 \\
      1 & 0 & 1 & 0 & 1 \\
      0 & 1 & 1 & 1 & 0
    \end{bmatrix}.
  \]
  Compute a realization of~$M^\star$, and some contractions of~$M$ of your choosing.

\end{enumerate}


\end{document}
